Dirac neutrino mass generation from Majorana messenger

### 2019-08-06 draft.tex
lc

The most general Lagrangian contains the following Yukawa and scalar interactions:
%
\begin{align*}
\label{Eq:LagY}
    \mathcal{L} \subset& -[ 
    h_{i\alpha} \overline{L_{i}} \tilde{\eta} \psi_{R\alpha} +  y_{j\alpha} \overline{\nu_{R_{j}}} \sigma^{\dagger} \psi^c_{R\alpha} + \kappa_{\alpha\beta} \overline{\psi^{c}_{R\alpha}} \psi_{R\beta} S^\dagger + \text{h.c.}] - \mathcal{V}(H, S, \eta, \sigma)\,,
\end{align*}
%
where $L_{i}$ and $H$ are the SM lepton and Higgs doublets, respectively,  $\widetilde{\eta} = i \sigma_2 \eta^{\dagger}$, and $h$, $y$ and $\kappa$ are matrices in the flavor space. 
The scalar potential can be cast as
%
\begin{align*}
    \mathcal{V}(H, S, \eta, \sigma) = & V(H) + V(S) + V(\eta) + V(\sigma) \\
    &+  \lambda_{1} (H^{\dagger} H ) (S^{*} S) + \lambda_{2} (H^{\dagger} H ) (\sigma^{*} \sigma ) + \lambda_{3} (H^{\dagger} H ) (\eta^{\dagger} \eta )\\
    &+ \lambda_{4} (S^{*} S) (\sigma^{*} \sigma ) + \lambda_{5} (S^{*} S) (\eta^{\dagger} \eta ) + \lambda_{6} (\eta^{\dagger} \eta ) (\sigma^{*} \sigma ) + \lambda_{7} (\eta^{\dagger} H ) (H^{\dagger} \eta ) \\
    &+ \lambda_{8} (\eta^{\dagger} H S^{*} \sigma + \text{h.c.})\,,
\end{align*}
%
with $V(\omega) = \mu^{2}_{\omega} \omega^{\dagger} \omega + \lambda_{\omega} (\omega^{\dagger} \omega)^{2}$. After the spontaneous symmetry breaking of $\operatorname{U}(1)_X$ the last term gives raise to the soft trilinear coupling
\begin{align}
  \mu=\lambda_8 \langle S \rangle\,.
\end{align}
We assume $\mu$ real to preserve CP symmetry in the scalar sector, and $\mu^2_\eta,\mu^2_\sigma>0$ to avoid tree-level mixing terms among the fermions. Moreover, the scalar couplings are subject to perturbativity and vacuum stability constraints:
\begin{align*}
\lambda_{H} &\geq 0\,, & \lambda_{\eta} &\geq 0\,, & \lambda_{S} &\geq 0\,, & \lambda_{\sigma} &\geq 0\,, & \lambda_{1} &\geq 0\,, & \\
\lambda_{2} &\geq 0\,, &
\lambda_{3}+\lambda_{7} &\geq 0\,, & \lambda_{4} &\geq 0\,, & \lambda_{5} &\geq 0\,, & \lambda_{6} &\geq 0\,,
\end{align*}
where we can set $\rho^{2} = 1$, because we assume that $\lambda_{7} \leq 0$.

@@

\section{Lepton flavor violation}
\label{sec:LFV}
%
\begin{figure}
\centering
\includegraphics[scale=0.6]{LFV.pdf}
\caption{Feynman diagram for process $l^{-}_{i} \to l^{-}_{j} \gamma$}
\label{fig:LFV}
\end{figure}
%
The new Yukawa interactions in Eq~\eqref{Eq:LagY} in our model introduce processes that violate the leptonic flavor in the charged sector (CLFV). These CLFV processes are induced at one loop level, as shown in the fig.~\ref{fig:LFV}. In this case, the $\overline{L} \tilde{\eta} \psi_R$ interaction allows the part of the double scalar $\eta$ ($\eta^{\pm}$) to mediate in the loop.

@@

\section{Dark matter}
\label{sec:DM}
Depending on the the new particles mass hierarchy (scalars and fermions charged odd under $U(1)_X$), it is possible to have two candidates for DM. The first DM candidate is the lightest majorana fermion (assuming that $\psi_{1} < \psi_{2} < \psi_{3}$). The second candidate is scalar DM. The scalar that assumes the role of DM depends on which is the lightest scalar particle charged odd under $U(1)_X$. In this model the scalar DM can be the neutral component of the scalar doublet $\eta^{0}$ or the scalar singlet $\sigma$.

\subsection{Fermionic dark matter}
In the case that the mass of the lightest fermion $\psi$ is smaller than the mass of the scalars, the DM will be fermionic. For this DM it is possible to have the following observable decay of the scalar $\eta^{\pm} \to l^{\pm} \psi_{1}$, therefore, the other fermions decay as $ \psi \to l^{\pm} l^{\mp} \psi_{1}$ and $\psi \to l^{\pm} l^{\mp} \psi_{1,2} $, through $\eta^{\pm}$. The relic density of DM can be controlled by annihilation $\psi_{1} \psi_{1} \to l _{\alpha} \overline{l}_{\beta} $, via Yukawa coupling $h^{a}_{i}$. These Yukawa couplings are restricted under LFV and neutrino masses. But, Yukawa couplings $y^{a}_{j}$ is free.

Because the particle $\psi_{1}$ only interacts with the leptons via the Yukawa links and this particle does not interact with gluons and quarks, this model has no direct detection restrictions for DM. In the case where the mass of the other fermions and scalars is nearly degenerated to $\psi_{1} $, the processes of coaniquilación are relevant in relic density~\cite{Klasen:2013jpa}.

\subsection{Scalar dark matter}
For scalar DM we have that the lightest scalar charged odd under $U(1)_X$ must have a mass smaller than the lightest $\psi$ fermion. Contrary to fermionic DM, scalar DM has additional interactions to Yukawa couplings (scalar and gauge interactions), which can be used to control relic density.

If the lightest scale is the neutral part of the scalar doublet, we will have that the DM will be similar to the model of the inert doublet~\cite{Honorez:2010re}. For the inert doublet model there are two mass regions for DM which can to reproduce adequately the relic density ($50$GeV \textless $m_{DM}$ \textless $70$GeV and $535$GeV \textless $m_{DM}$ \textless $20$TeV)~\cite{Garcia-Cely:2015khw}. In recent studies, these two regions are connected through use of a second doublet~\cite{Borah:2019aeq}. How we have in the model the Yukawa coupling, the decay $\psi_{i} \to l^{\pm} \eta^{\mp}$ is possible. In addition, decay is given via scalar interactions $\eta^{\pm} \to \eta^{0} + {W^{\pm}}$, where $W^{\pm}$ can be real or virtual and decay to leptons or quarks.

Now, if the lightest scale is the scalar singlet, the DM will be similar to the model in Ref.~\cite{Athron:2017kgt}. The mass range for the scalar singlet that produces the adequate relic density is between approximately the Higgs mass and about $300$GeV and masses above $1$TeV. For this model, the term in the Lagrangian $\sigma^{2} H^{2}_{i}$ is known as the \textquotedblleft Higgs portal\textquotedblright. This term generates interesting phenomenological consequences. Such as, thermal production of DM in the early universe~\cite{Yaguna:2008hd}, direct detection and $h \to \sigma \sigma$~\cite{Mambrini:2011ik}. In addition, this model has important contributions to cosmology, in phenomena such as inflation~\cite{Lerner:2009xg} and baryogenesis~\cite{Cline:2012hg}.

@@

%
\begin{table}
  \centering
  \begin{tabular}{|l|l|c|c|}
    \hline  
    Particles     & $\left( SU(2)_L, U(1)_Y \right)$ & $U(1)_{L}$& $Z_2$\\ \hline
    $H $  & $(2, 1/2)$ &  0 &$+$\\
    %$\widetilde{H_{2}}$  & $(2,1/2)$ &  $0$ &$-$\\
    $\eta$ & $(2,1/2)$ & $1-\psi\to 5/2$ &$-$\\
    $S$ & $(1,0)$ & $-2\psi\to 3$ &$+$\\
    $\sigma$ & $(1,0)$ & $\nu+\psi=\eta=1-\psi\to 5/2$&$-$\\
    \hline
    $Q_{L_{i}}$  & $(2,-1/6)$ & $0$& $+$\\
    $\overline{u_{R_{i}}}$ & $(1,-2/3)$ & $0$& $+$\\
    $\overline{d_{R_{i}}}$ & $(1,1/3)$ & $0$& $+$\\
    \hline
    $L_i$  & $(2,-1/2)$ & $-1$& $+$\\
    $\overline{e_{R_i}}$ & $(1,1)$ & $+1$ & $+$ \\
    $\overline{\nu_{R_{i,j}}}$ & $(1,0)$ & $\nu\to 1-2\psi\to 4$&$+$\\
    $\overline{\nu_{R_k}}$ & $(1,0)$ & $\nu_k$&$+$\\
    $\psi_{i}$  & $(1,0)$ & $\psi\to -3/2$ &$-$\\\hline
  \end{tabular}
  \caption{Scalars and left-handed fermions with their respective charges. $Z_2$ to define THDM Type-II~\cite{Chiang:2013ixa}}
  \label{tab:partcont}
\end{table}
%

% Checl also:
$-2,6,-9,-3,7$

>>>>

The most general Lagrangian contains the following Yukawa and scalar interactions:
%
\begin{align*}
\label{eq:LagY}
    \mathcal{L} \subset& -[ 
    h_{i\alpha} \overline{L_{i}} \tilde{\eta} \psi_{R\alpha} +  y_{j\alpha} \overline{\nu_{R_{j}}} \sigma^{\dagger} \psi^c_{R\alpha} + \kappa_{\alpha\beta} \overline{\psi^{c}_{R\alpha}} \psi_{R\beta} S^\dagger + \text{h.c.}] - \mathcal{V}(H, S, \eta, \sigma)\,,
\end{align*}
%
where $L_{i}$ and $H$ are the SM lepton and Higgs doublets, respectively,  $\widetilde{\eta} = i \sigma_2 \eta^{\dagger}$, and $h$, $y$ and $\kappa$ are matrices in the flavor space. 
The scalar potential can be cast as
%
\begin{align*}
    \mathcal{V}(H, S, \eta, \sigma) = & V(H) + V(S) + V(\eta) + V(\sigma) \\
    &+  \lambda_{1} (H^{\dagger} H ) (S^{*} S) + \lambda_{2} (H^{\dagger} H ) (\sigma^{*} \sigma ) + \lambda_{3} (H^{\dagger} H ) (\eta^{\dagger} \eta )\\
    &+ \lambda_{4} (S^{*} S) (\sigma^{*} \sigma ) + \lambda_{5} (S^{*} S) (\eta^{\dagger} \eta ) + \lambda_{6} (\eta^{\dagger} \eta ) (\sigma^{*} \sigma ) + \lambda_{7} (\eta^{\dagger} H ) (H^{\dagger} \eta ) \\
    &+ \lambda_{8} (\eta^{\dagger} H S^{*} \sigma + \text{h.c.})\,,
\end{align*}
%
with $V(\omega) = \mu^{2}_{\omega} \omega^{\dagger} \omega + \lambda_{\omega} (\omega^{\dagger} \omega)^{2}$. After the spontaneous symmetry breaking of $\operatorname{U}(1)_X$ the last term gives raise to the soft trilinear coupling
\begin{align}
  \mu=\lambda_8 \langle S \rangle\,.
\end{align}
We assume $\mu$ real to preserve CP symmetry in the scalar sector, and $\mu^2_\eta,\mu^2_\sigma>0$ to avoid tree-level mixing terms among the fermions. Moreover, the scalar couplings are subject to perturbativity and vacuum stability constraints:
\begin{align*}
\lambda_{H} &\geq 0\,, & \lambda_{\eta} &\geq 0\,, & \lambda_{S} &\geq 0\,, & \lambda_{\sigma} &\geq 0\,, & \lambda_{1} &\geq 0\,, & \\
\lambda_{2} &\geq 0\,, &
\lambda_{3}+\lambda_{7} &\geq 0\,, & \lambda_{4} &\geq 0\,, & \lambda_{5} &\geq 0\,, & \lambda_{6} &\geq 0\,,
\end{align*}
where we can set $\rho^{2} = 1$, because we assume that $\lambda_{7} \leq 0$.

@@

\section{Lepton flavor violation}
\label{sec:LFV}
%
\begin{figure}
\centering
\includegraphics[scale=0.6]{LFV.pdf}
\caption{Feynman diagram for process $l^{-}_{i} \to l^{-}_{j} \gamma$}
\label{fig:LFV}
\end{figure}
%
The new Yukawa interactions in eq.~\eqref{eq:LagY} in our model introduce processes of charged lepton flavor violation (CLFV). These CLFV processes are induced at one loop level, as shown in the fig.~\ref{fig:LFV}. In this case, the $\overline{L} \tilde{\eta} \psi_R$ interaction allows the part of the double scalar $\eta$ ($\eta^{\pm}$) to mediate in the loop.

@@

\section{Dark matter}
\label{sec:DM}
From table~\ref{tab:partcont}. we can see that after the spontaneous breaking of
$\operatorname{U}(1)_X$ there is remnant $Z_2$ symmetry under which the particles circulating the loop in fig.~\ref{fig:zee}. plus $\nu_{R3}$ are odd, while the other right handed neutrinos and the SM particles are even. The lightest odd particle can be a dark matter candidate. In this way we can have either fermionic or scalar dark matter in the model.

\subsection{Fermionic dark matter}
The relic density of DM is controlled by annihilation $\psi_{1} \psi_{1} \to l _{\alpha}^+ \overline{l}_{\beta}^- $, via Yukawa coupling $h^{a}_{i}$. These Yukawa couplings are restricted by CLFV and neutrino masses. %But, Yukawa couplings $y^{a}_{j}$ is free.

Because the particle $\psi_{1}$ only interacts with the leptons via the Yukawa couplings, this model has no a direct detection cross section at three level. In the case where the mass of the other fermions and scalars is nearly degenerated to $\psi_{1} $, the processes of coannihilation  can be relevant for the relic density calculation~\cite{Klasen:2013jpa}.

% In the case that the mass of the lightest fermion $\psi$ is smaller
% than the mass of the scalars, the DM will be fermionic. For this DM it
% is possible to have the following observable decay of the scalar
% $\eta^{\pm} \to l^{\pm} \psi_{1}$, therefore, the other fermions decay
% as $ \psi \to l^{\pm} l^{\mp} \psi_{1}$ and
% $\psi \to l^{\pm} l^{\mp} \psi_{1,2} $, through $\eta^{\pm}$.

\subsection{Scalar dark matter}
Contrary to fermionic DM, scalar DM has additional interactions beyond the Yukawa couplings  which can be used to control relic density. This corresponds to the simplified dark matter model of scalar singlet-doblet dark matter, which have been studied in detail in~\cite{}

When the lightest scalar  is mainly a scalar doublet, the phenomenology will be similar inert doublet model~\cite{Honorez:2010re}. For this kind of DM model  inert doublet model there are two mass regions for DM which can to reproduce adequately the relic density ($50$GeV \textless $m_{DM}$ \textless $70$GeV and $535$GeV \textless $m_{DM}$ \textless $20$TeV)~\cite{Garcia-Cely:2015khw}.

% In recent studies, these two regions are connected through use of a
% second doublet~\cite{Borah:2019aeq}. How we have in the model the
% Yukawa coupling, the decay $\psi_{i} \to l^{\pm} \eta^{\mp}$ is
% possible. In addition, decay is given via scalar interactions
% $\eta^{\pm} \to \eta^{0} + {W^{\pm}}$, where $W^{\pm}$ can be real or
% virtual and decay to leptons or quarks.

Now, if the lightest scalar is mainly a singlet, the DM will be
similar to Singlet Scalar Dark Matter
Model~\cite{Athron:2017kgt}\todo{change for the original ref by Ma et
  al. etc}.
The mass range for the scalar singlet that produces the adequate relic
density is around the Higgs resonance and for masses above $1$TeV. For
this model~\cite{}, the relevant coupling for DM phenomenology is the
Higgs portal trough the Lagrangian term $\sigma^{2} H^{2}_{i}$.
This term generates interesting phenomenological consequences.
Such as, thermal production of DM in the early
universe~\cite{Yaguna:2008hd}, direct detection and invisible decay of
SM Higgs through $h \to \sigma \sigma$~\cite{Mambrini:2011ik}.
In addition, this model has important contributions for cosmology, in
phenomena such as inflation~\cite{Lerner:2009xg} and
baryogenesis~\cite{Cline:2012hg}.

@@

%
\begin{table}
  \centering
  \begin{tabular}{|l|l|c|c|}
    \hline  
    Particles     & $\left( SU(2)_L, U(1)_Y \right)$ & $U(1)_{L}$& $Z_2$\\ \hline
    $H $  & $(2, 1/2)$ &  0 &$+$\\
    $\eta$ & $(2,1/2)$ & $1-\psi\to 5/2$ &$-$\\
    $S$ & $(1,0)$ & $-2\psi\to 3$ &$+$\\
    $\sigma$ & $(1,0)$ & $\nu+\psi=\eta=1-\psi\to 5/2$&$-$\\
    $L_i$  & $(2,-1/2)$ & $-1$& $+$\\
    $\overline{e_{R_i}}$ & $(1,1)$ & $+1$ & $+$ \\
    $\overline{\nu_{R_{i,j}}}$ & $(1,0)$ & $\nu\to 1-2\psi\to 4$&$+$\\
    $\overline{\nu_{R_k}}$ & $(1,0)$ & $\nu_k$&$+$\\
    $\psi_{i}$  & $(1,0)$ & $\psi\to -3/2$ &$-$\\\hline
  \end{tabular}
  \caption{Scalars and left-handed fermions with their respective charges.}
  \label{tab:partcont}
\end{table}
%
